\thesistitle

### 2019-07-25 sections/clonerefactor.tex
Removed experimental setup and merged it with clonerefactor

\subsubsection{Try to Expand}
The problem with the graph clone detection technique is that

@@

>>>>

\subsubsection{Try to Expand}
The problem with the graph clone detection technique is that ...
\todo{TODO}

@@

\section{Context Analysis of Clones}\label{chap:contextsetup}
To be able to refactor code clones, CloneRefactor maps the context of code clones. We define the following aspects of the clone as its context:
\begin{enumerate}
  \item \textbf{Relation:} The relation of clone instances among each other through inheritance.
  \item \textbf{Location:} Where a clone instance occurs in the code.
  \item \textbf{Contents:} The statements/declarations of a clone instance.
\end{enumerate}
We define categories for each of these aspects and enable CloneRefactor to determine the catagories of clones.

\begin{figure}[H]
  \centering
    \includegraphics[width=0.8\columnwidth]{img/context}
    \caption{Abstract representation of clone classes and clone instances.}
  \label{fig:clonecontext}
\end{figure}

Fig.~\ref{fig:clonecontext} shows an abstract representation of clone classes and clone instances. The relation of clones through inheritance is measured for each clone class. The location and contents of clones are measured for each clone instance.

\subsection{Relation}\label{sec:setuprelation}
When merging code clones in object-oriented languages, it is important to consider the inheritance relation between clone instances. This relation has a big impact on how a clone should be refactored.

Fontana et al.~\cite{fontana2015duplicated} describe measurements on 50 open source projects on the relation of clone instances to each other. To do this, they first define several categories for the relation between clone instances in object-oriented languages. These categories are as follows:
\begin{enumerate}
  \item \textbf{Same Method}: All instances of the clone class are in the same method.
  \item \textbf{Same Class}: All instances of the clone class are in the same class.
  \item \textbf{Superclass}: All instances of the clone class are in a class that is child or parent of each other.
  \item \textbf{Sibling Class}: All instances of the clone class have the same parent class.
    \item \textbf{Ancestor Class}: All instances of the clone class are superclasses except for the direct superclass.
  \item \textbf{First Cousin Class}: All instances of the clone class have the same grandparent class.
\item \textbf{Same Hierarchy Class}: All instances of the clone class belong to the same inheritance hierarchy, but do not belong to any of the other categories.
\item \textbf{Same External Superclass}: All instances of the clone class have the same superclass, but this superclass is not included in the project but part of a library.
\item \textbf{Unrelated class}: There is at least one instance in the clone class that is not in the same hierarchy.
\end{enumerate}

We added the following categories, to gain more information about clones and reduce the number of unrelated clones:

\begin{enumerate}
\item \textbf{Same Interface}: All instances of the clone class are in a class or interface that have a common interface anywhere in their inheritance hierarchy.
\item \textbf{No Direct Superclass}: All instances of the clone class are in a class that does not have any superclass.
\item \textbf{No Indirect Superclass}: All instances of the clone class are in a class that does not have any external classes in its inheritance hierarchy.
\item \textbf{External Ancestor}: All instances of the clone class are in a class that does not have any external classes in its inheritance hierarchy.
\end{enumerate}

Figure \ref{fig:clonecontext} shows an abstract representation of clone classes and clone instances. The relation of clones through inheritance is measured on clone class level: it involves all child clone instances. The location and contents of clones is measured on clone instance level. A clone's location involves the file it resides in and the range it spans (for example: line 6 col 2 - line 7 col 50). A clone instance contents consists of a list of all statements and declarations it spans.

\begin{figure}[H]
  \centering
    \includegraphics[width=0.6\columnwidth]{img/Relation}
      \caption{Abstract figure displaying relations of clone classes. Arrows represent superclass relations.}
  \label{fig:clonerelation}
\end{figure}

We separate these relations into the following categories, because of their related refactoring opportunities:
\begin{itemize}
  \item \textbf{Common Class}: \textit{Same Method}, \textit{Same Class}
  \item \textbf{Common Hierarchy}: \textit{Superclass}, \textit{Sibling Class}, \textit{Ancestor Class}, \textit{First Cousin}, \textit{Same Hierarchy}
  \item \textbf{Common Interface}: \textit{Same Interface}
  \item \textbf{Unrelated}: \textit{No Direct Superclass}, \textit{No Indirect Superclass}, \textit{External Superclass}, \textit{External Ancestor}
\end{itemize}

Every clone class only has a single relation, which is the first relation from above list that the clone class applies to. For instance: all ``Superclass'' clones also apply to ``Same Hierarchy'', but because ``Superclass'' is earlier in above list they will get the ``Superclass'' relation. This is because the items earlier in the list denote a more favorable refactoring.

To check a the relation for a given clone class, we compare each node in a clone instance with the respective nodes in each other clone instance within a clone class. So for a clone class with x clone instances and each clone instance has y nodes, we perform:

\begin{equation}\label{eq:sameclass}
\frac{x * (x + 1)}{2} * y
\end{equation}

comparisons to determine the relation. Each of these checks will result in one of the relations from the list above. Of these relations, the item lowest in the list is going to be the relation of the clone class. So for instance, if a clone class has 15 nodes that denote a \textit{Superclass} relation but 3 nodes are \textit{Unrelated}, the clone class becomes \textit{Unrelated}.

\subsubsection{Common Class}
The \textit{Same method} and \textit{Same class} relations share a common refactoring opportunity. Clones of both these catagories, when extracted to a new method, can be placed in the same class. Both of these relations are most favorable for refactoring, as they require a minimal design tradeoff. Furthermore, global variables that are used in the class can be used without having to create method parameters.

Cloned nodes are flagged as \textit{Same method} if these nodes are found in the same method. We define the method as the first method declaration that is encountered when following the parent nodes in the ast for a given cloned node. Please note that a clone instance may not always be in a method, for which this predicate will fail.

Cloned nodes are flagged as \textit{Same class} if these nodes are found in the same class. We define the class as the first class declaration that is encountered when following the parent nodes in the ast for a given cloned node.

\subsubsection{Common Hierarchy}
Clones that are in a common hierarchy can be refactored by using the ``Extract Method'' refactoring method followed by ``Pull Up Method'' until the method reaches a location that is accessible by all clone instances. However, the more often ``Pull Up Method'' has to be used, the more detrimental the effect is on system design. This is because putting a lot of functionality in classes higher up in an inheritance structure can result in the ``God Object'' anti-pattern. A god object is an object that knows too much or does too much.

Cloned nodes are flagged as \textit{Superclass} if the classes in which these nodes are found are parent and child class of each other. Cloned nodes are flagged as \textit{Siblings} if the classes in which these nodes are found all share the same parent and this parent class is not external. Cloned nodes are flagged as \textit{First Cousin} if the classes in which these nodes are found all share the same grandparent and this grandparent class is not external.

Cloned nodes are flagged as \textit{Ancestor} if the classes in which these nodes are found are recursively the parent of an antecedent. Cloned nodes are flagged as \textit{Same Hierarchy} if the classes in which these nodes are found are all in the same inheritance hierarchy and not linked by any external classes.

\subsubsection{Common Interface}
Many object-oriented languages know the concept of ``interfaces'', which are used to specify a behavior that classes must implement. As code clones describe functionality and interfaces originally did not allow for functionality, interfaces did not open up refactoring opportunities for duplicated code. However, many programming languages nowadays support default implementations in interfaces. Since Java 7 and C\#8, these programming languages allow for functionality to be defined in interfaces. Many other object-oriented languages like Python allow this by nature, as they do not have a true notion of interfaces.

The greatest downside on system design of putting functionality in interfaces is that interfaces are per definition part of a classes' public contract. That is, all functionality that is shared between classes via an interface cannot be hidden by setting a stricter visibility. Because of that, we favor all ``Common Hierarchy'' refactoring opportunities over ``Common Interface''.

To check whether two cloned nodes have common interfaces, we recursively walk all types the class or interface of the cloned nodes implement and extend. The common interface that is closest to the cloned node in terms of depth of recursion will be flagged as the refactoring candidate.

\subsubsection{Unrelated}
Clones are unrelated if they share no common class or interface in their inheritance structure. These clones are least favorable when looking at refactoring, because their refactoring will almost always have a major impact on system design. We formulated four catagories of unrelated clones to look into their refactoring opportunities.

Cloned nodes are flagged \textit{No Direct Superclass} if they are in classes that do currently not have a parent. This marks the opportunity for creating a superclass abstraction and placing the extracted method there. Cloned nodes are flagged \textit{No Indirect Superclass} if any of their ancestors does not have a parent. In such a case, it would be possible to create such an abstraction for the ancestor that does not have a parent.

Cloned nodes are flagged \textit{External Superclass} if they are in a class which has an external parent. Cloned nodes are flagged \textit{External Ancestor} if one of their ancestors has an external parent. Both of these relations obstruct the possibility of creating a superclass abstraction. In such a case, an interface abstraction could be created to make their relation explicit.

\subsection{Location}\label{sec:setuplocation}
A paper by Lozano et al. \cite{lozano2007evaluating} discusses the harmfulness of cloning. The authors argue that 98\% are produced at method-level. However, this claim is based on a small dataset and based on human copy-paste behavior rather than static code analysis. We decided to measure the locations of clones through static analysis on our dataset. We chose the following categories:
\begin{enumerate}
  \item \textbf{Method/Constructor Level:} A clone instance that does not exceed the boundaries of a single method or constructor (optionally including the declaration of the method or constructor itself).
  \item \textbf{Class Level:} A clone instance in a class, that exceeds the boundaries of a single method or contains something else in the class (like field declarations, other methods, etc.).
  \item \textbf{Interface Level:} A clone that is (a part of) an interface.
  \item \textbf{Enumeration Level:} A clone that is (a part of) an enumeration.
\end{enumerate}
We check the location of each clone instance for each of its nodes. If any node reports a different location from the others, we choose the location that is lowest in above list. So for instance, if a clone instance has 15 nodes that denote a \textit{Method Level} location but 3 nodes are \textit{Class Level}, the clone instance becomes \textit{Class Level}.

\todo{Explain why these categories indicate and how they are related to refactoring}

\subsection{Contents}\label{sec:setupcontents}
Finally, we looked at what nodes individual clone instances span. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans a method partially, optionally including its declaration.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Includes Fields/Constructor:} A clone that spans a combination of fields and other things, like methods.
  \item \textbf{Method/Class/Interface/Enumeration Declaration:} A clone that contains the declaration (usually the first line) of a class, method, interface or enumeration.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

\todo{Explain each category in more detail}

\section{Method extraction opportunities}\label{sec:refactorabilitysetup}
The most used technique to refactor clones is method extraction (creating a new method on basis of the contents of clones). However, method extraction cannot be applied in all cases. In these instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all.

We measured the number of clones that can be refactored through method extraction (without additional transformations being required). We defined the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table~\ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}

\todo{Explain each category in more detail}
